---
title: "Assignment 5"
author: "Michael D'Acampora"
date: "October 1, 2017"
output: html_document
---
```{r}
library(tidyr)
library(dplyr)

flt <- read.csv("Assignment5.csv", na.strings = "")
flt
```

Use tidyr and dplyr to clean up the table in one fell swoop, and add a column showing on-time performance percentage.
```{r}
tidy.flt <- flt %>% 
  
  drop_na(Los.Angeles) %>% 
  
  rename("Airline"=ï.., "Status"=X) %>% 

  fill(Airline) %>% 
  
  gather(City, No.Flights, Los.Angeles:Seattle) %>%
  
  spread(Status, No.Flights) %>% 
  
  rename("Delayed"=delayed, "On.Time"= "on time") %>% 
  
  mutate(PercentOn.Time = round(
                              (On.Time / (On.Time + Delayed))*100,1))

tidy.flt

```

See total flights for each airline
```{r}
tidy.flt %>% 
  group_by(Airline) %>%  
  
  summarise(Total.Flights = sum(Delayed) + sum(On.Time))

```

See total on-time flights for each airline
```{r}
tidy.flt %>% 
  group_by(Airline) %>% 
  
  summarise(On.Time = sum(On.Time))


```

See total delayed flights for each airline
```{r}
tidy.flt %>% 
  group_by(Airline) %>% 
  
  summarise(Delayed = sum(Delayed))


```

Grab the weighted average of on-time performance to see which airline has been better overall.

At first I only averaged the OnTimePerformance percentages by City and Alaska Airlines actually looked like it had better on-time performance. However, since weighted average is better suited and since AM West was so strong and voluminous in Phoenix (92.1%), it pulled their weighted average up. 

It also appears that Alaska Airlines had one of their worst on-time performances in Seattle, which had more than triple the number of flights compared to the other citys where the company flew.
```{r}
tidy.flt %>% 
  group_by(Airline) %>% 
  
  summarise(AvgPercentOnTime = round(sum(On.Time) / sum(sum(On.Time)+sum(Delayed))*100,1))
```

