---
title: "Project 1 - Chess Tournament Table"
author: "Michael D'Acampora"
date: "September 24, 2017"
output: html_document
---

```{r}
library(stringr)

data <- "tournamentinfo.txt"

dat <- readLines(data)

head(dat, 33)
```

Split the file into two tables by odd and even rows with the sequence function.
```{r}
chessA <- dat[seq(5,195,3)]
head(chessA, 10)
```

```{r}
chessB <- dat[seq(6,196,3)]
head(chessB, 10)
```

Extract the name using str_extract_all on the odd rows. Use unlist to vectorize the output and str_trim to clean up the empty space.
```{r}
name <- str_trim(unlist(str_extract_all(chessA,'[A-Z]+ [A-Z]+ ([A-Z-]+)? ([A-Z]+)?')))
head(name, 10)
```

Extract the total pts using str_extract_all on the odd rows. Use unlist to vectorize the output and make it numeric.
```{r}
total <- as.numeric(unlist(str_extract_all(chessA, "\\d\\.\\d")))
head(total, 10)
```

Extract the state using str_extract_all on the even rows. Use unlist to vectorize the output and use str_trim to clean up the empty space.
```{r}
state <- str_trim(unlist(str_extract_all(chessB, "\\s[A-Z]{2}\\ ")))
head(state, 10)
```

Extract the player prerating using str_extract_all. Use unlist to vectorize the output.
Complete a second pass of clean up and make it numeric. 

```{r}
prerate <- unlist(str_extract_all(chessB, ":\\s+\\d{3,4}"))
prerate <- as.numeric(unlist(str_extract_all(prerate, "\\d{1,4}")))
head(prerate, 10)
```

Describing the for loop:

The tables have 64 rows. For each row:

step 1: extract the W,D, or L and opponent number from each row 

step 2: clean up the W,D, or L and spacing using str_replace, and make number numeric

step 3: extract the player rating of the row number assigned to each of the values of step 2

step 4: clean up the extraction with str_replace_all and make opponent ratings numeric

step 5: sum up the row of opponent ratings and divide by the amount of opponents

```{r}

opp.rate <- vector()

for (i in 1:64){
  
  s1 <- unlist(str_extract_all(chessA[i], "(W|D|L)\\s+(\\d)+"))
  
  s2 <- as.numeric(str_replace_all(s1, "(W|D|L)\\s+", ""))
  
  s3 <- str_extract(chessB[s2[1:length(s2)]], ":\\s+\\d{3,4}")
  
  s4 <- as.numeric(str_replace_all(s3, ":\\s+", ""))
  
  s5 <- round(sum(s4)/length(s4))
  
  opp.rate <- c(opp.rate, s5)
}

head(opp.rate, 10)
```

Create a data frame with the five columns created above
```{r}

chess_df <- data.frame(
                      PlayerName = name, 
                      State = state, 
                      TotalPts = total, 
                      AvgPrerating = prerate,                         
                      OppAvgPrerating = opp.rate
            )

chess_df
```

Write to csv and remove the row numbers
```{r}
write.csv(chess_df, file = "tournamenttable.csv", row.names = FALSE)
```



