---
title: "Project 2 Dataset 2"
author: "Michael D'Acampora"
date: "October 7, 2017"
output: html_document
---

### This dataset is based on approximately the last 5 years of NYPD motor vehicle collisions in New York City. 
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)

data2 <- read.csv("NYPD_Motor_Vehicle_Collisions.csv")
head(data2)

```

#### checking the data types
```{r}
sapply(data2, mode)
sapply(data2, class)
```



#### select the relvant variables, clean up the column names 
```{r}
Accidents <- data2 %>% 
  select(DATE,
         TIME,
         BOROUGH, 
         ZIP.CODE,
         NUMBER.OF.PERSONS.INJURED,
         NUMBER.OF.PERSONS.KILLED,
         NUMBER.OF.PEDESTRIANS.INJURED,
         NUMBER.OF.PEDESTRIANS.KILLED,
         NUMBER.OF.CYCLIST.INJURED,
         NUMBER.OF.CYCLIST.KILLED,
         NUMBER.OF.MOTORIST.INJURED,
         NUMBER.OF.MOTORIST.KILLED,
         CONTRIBUTING.FACTOR.VEHICLE.1,
         CONTRIBUTING.FACTOR.VEHICLE.2,
         VEHICLE.TYPE.CODE.1,
         VEHICLE.TYPE.CODE.2) %>% 

  rename('PERS.INJ'    = NUMBER.OF.PERSONS.INJURED,
         'PERS.KILLED' = NUMBER.OF.PERSONS.KILLED,
         'PED.INJ'     = NUMBER.OF.PEDESTRIANS.INJURED,
         'PED.KILLED'  = NUMBER.OF.PEDESTRIANS.KILLED,
         'CYC.INJ'     = NUMBER.OF.CYCLIST.INJURED,
         'CYC.KILLED'  = NUMBER.OF.CYCLIST.KILLED,
         'MOT.INJ'     = NUMBER.OF.MOTORIST.INJURED,
         'MOT.KILLED'  = NUMBER.OF.MOTORIST.KILLED)

head(Accidents)

```


#### Build a new table grouped by borough
```{r}
By.Borough <- Accidents %>% 
  select(BOROUGH, 
         PERS.INJ,
         PERS.KILLED,
         PED.INJ,
         PED.KILLED,
         CYC.INJ,
         CYC.KILLED,
         MOT.INJ,
         MOT.KILLED) %>% 
  
  group_by(BOROUGH) %>% 
  
  summarise(TOTAL.INJ    = sum(PERS.INJ),
            TOTAL.KILLED = sum(PERS.KILLED),
            PED.INJ      = sum(PED.INJ),
            PED.KILLED   = sum(PED.KILLED),
            CYC.INJ      = sum(CYC.INJ),
            CYC.KILLED   = sum(CYC.KILLED),
            MOT.INJ      = sum(MOT.INJ),
            MOT.KILLED   = sum(MOT.KILLED)) 
  

By.Borough
```



#### Plot the total killed by borough. There is a huge contingent of empty borough data. The zipcode data was also empty. It seems the onlyi thing you could do is figure out a way to map the GPS data to the relevant borough to fill in the raw data. 
```{r}
ggplot(By.Borough,mapping = aes(BOROUGH, TOTAL.KILLED)) +
  geom_bar(stat = 'identity')

```



#### Compare total injured by zip code, drop the NA rows
```{r}
By.Zip <- Accidents %>% 
  select(ZIP.CODE,
         PERS.INJ,
         PERS.KILLED) %>% 
  
  group_by(ZIP.CODE) %>% 
  
  summarise(Total.Injured = sum(PERS.INJ),
            Total.Killed  = sum(PERS.KILLED)) %>% 
  
  drop_na(ZIP.CODE)

# get summary data for total injured and total killed
summary(By.Zip$Total.Injured)
summary(By.Zip$Total.Killed)

```



