---
title: "Project 2 - Dataset 3"
author: "Michael D'Acampora"
date: "October 7, 2017"
output: html_document
---

### This dataset #3 is 25 years worth of catfish stocks
```{r}

library(tidyr)
library(dplyr)
library(ggplot2)

data3 <- read.csv("CatfishFarm.csv", header=FALSE)
head(data3)
```

#### Drop the first row, and tidy by using rename and gather. There was an issue with this data set regarding column names and I felt this was the best way to do it.
```{r}
data3 <- data3[-c(1), ]

# tidy up: rename and gather
catfish <- data3 %>% 
  rename("Category"=V1,
         "1992"=V2,
         "1993"=V3,
         "1994"=V4,
         "1995"=V5,
         "1996"=V6,
         "1997"=V7,
         "1998"=V8,
         "1999"=V9,
         "2000"=V10,
         "2001"=V11,
         "2002"=V12,
         "2003"=V13,
         "2004"=V14,
         "2005"=V15,
         "2006"=V16,
         "2007"=V17,
         "2008"=V18,
         "2009"=V19,
         "2010"=V20,
         "2011"=V21,
         "2012"=V22,
         "2013"=V23,
         "2014"=V24,
         "2015"=V25,
         "2016"=V26) %>% 
  
  gather(-Category, key = "Year", value = "Stock")

head(catfish)
```

#### remove the commas from Stock so they will sum together, and compare stock by size/category
```{r}

catfish$Category <- gsub("/","", catfish$Category)
catfish$Stock <- as.numeric(gsub(",", "", catfish$Stock))

Stock.by.Size <- catfish %>% 
  
  select(Category, Stock) %>% 
  
  group_by(Category) %>% 
  
  summarise(Stock.by.Category = sum(Stock))

head(Stock.by.Size)
```

####  compare stock by year
```{r}
Stock.by.year <- catfish %>% 
  
  select(Year, Stock) %>% 
  
  group_by(Year) %>% 
  
  summarise(Stock = sum(Stock))

head(Stock.by.year)
```



#### Plot the stock by year. As you can see, catfish stockpiles have been on a downward trend since 2008. Could there be a connection with the 2008 financial crisis? Or maybe overfishing is making catfish a more popular item on the menu?
```{r}
ggplot(Stock.by.year,mapping = aes(Year, Stock)) +
  geom_bar(stat = 'identity')
```





